---
title: 'Operation Silent Web: Cybersecurity Threat Analysis'
author: 'Christopher Davis'
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
# Load necessary libraries
library(tidyverse)
library(lubridate)
library(knitr)
library(forecast)
```

# Executive Summary

This analysis simulates an FBI-led cybersecurity investigation using a mock dataset of over 1,000 cyber incidents reported across the United States between 2023 and 2024. The investigation identifies **phishing and ransomware** as the most frequent and damaging attack types, disproportionately affecting the **healthcare and banking sectors**. Geographically, **California and New York** emerged as primary hotspots. A time-series forecast projects a stable monthly incident volume of approximately **40–42 attacks in 2025**, with a confidence range between **34 and 50 incidents**. These findings suggest a continued threat environment that, while steady, still requires strategic readiness, sector-specific defenses, and enhanced interagency coordination.

---

# 1. Data Overview

```{r data-load}
# Load the dataset
cyber_data <- read_csv("data/cyber_incidents_mock.csv")
cyber_data$incident_date <- as.Date(cyber_data$incident_date)

# Display the structure of the data
glimpse(cyber_data)
```

This dataset includes 1,000 simulated cyberattack records with details such as attack type, severity, targeted sector, state of occurrence, incident date, and estimated financial loss.

---

# 2. Attack Type Frequency

```{r attack-type-plot}
cyber_data %>%
  count(attack_type) %>%
  ggplot(aes(x = reorder(attack_type, n), y = n)) +
  geom_col(fill = "steelblue") +
  coord_flip() +
  labs(
    title = "Attack Type Frequency",
    x = "Attack Type",
    y = "Number of Incidents"
  )
```

**Phishing** is the most common cyberattack type, followed by **ransomware** and **malware**. These categories make up the majority of reported threats and are particularly damaging in sectors handling sensitive or financial data.

---

# 3. Severity Levels

```{r severity-plot}
cyber_data %>%
  count(severity_level) %>%
  ggplot(aes(x = severity_level, y = n)) +
  geom_col(fill = "firebrick") +
  labs(
    title = "Severity Level Distribution",
    x = "Severity Level",
    y = "Number of Incidents"
  )
```

Most incidents are categorized as **Low** or **Medium** severity. However, a notable percentage of **Critical** cases warrants resource prioritization for early detection and response.

---

# 4. Industry Sector Impact

```{r sector-impact}
cyber_data %>%
  count(target_sector) %>%
  ggplot(aes(x = reorder(target_sector, n), y = n)) +
  geom_col(fill = "darkorange") +
  coord_flip() +
  labs(
    title = "Incident Volume by Sector",
    x = "Sector",
    y = "Incident Count"
  )
```

The **Healthcare** and **Banking** sectors were the most frequently targeted, aligning with national patterns where attackers seek maximum disruption or financial gain.

---

# 5. Geographic Hotspots

```{r hotspot-map}
cyber_data %>%
  count(reported_location) %>%
  top_n(5) %>%
  ggplot(aes(x = reorder(reported_location, n), y = n)) +
  geom_col(fill = "purple") +
  coord_flip() +
  labs(
    title = "Top 5 States by Incident Volume",
    x = "State",
    y = "Incidents"
  )
```

States like **California, New York, Texas, and Florida** lead in reported incidents, signaling high-value targets or more aggressive threat exposure.

---

# 6. Incident Trends Over Time

```{r trend-line}
cyber_data %>%
  mutate(month = floor_date(incident_date, "month")) %>%
  count(month) %>%
  ggplot(aes(x = month, y = n)) +
  geom_line(color = "darkgreen", size = 1) +
  geom_smooth(se = FALSE, color = "black") +
  labs(
    title = "Monthly Incident Trend",
    x = "Month",
    y = "Incidents"
  )
```

The trend shows a **steady volume of cyber incidents** over time, with minor fluctuations that may correspond to seasonal behaviors or policy changes.

---

# 7. Forecast for 2025

```{r forecast-model}
# Aggregate incidents by month
monthly_data <- cyber_data %>%
  mutate(month = floor_date(incident_date, "month")) %>%
  count(month)

# Create time series
incident_ts <- ts(monthly_data$n, start = c(2023, 1), frequency = 12)

# Forecast
model <- auto.arima(incident_ts)
forecasted <- forecast(model, h = 12)

# Plot forecast
autoplot(forecasted) +
  labs(
    title = "Forecasted Monthly Cyber Incidents (2025)",
    x = "Month",
    y = "Predicted Incident Count"
  )
```

Based on time-series modeling, the monthly incident volume is expected to average **40–42**, with a 95% confidence range of **34 to 50 incidents**. This stability masks the continued need for constant vigilance.

---

# 8. Recommendations

- Prioritize cyber defense investments in **healthcare and banking sectors**
- Deploy additional monitoring and field assets in **California, New York, Texas**
- Launch early-warning initiatives during historically active months
- Sustain 24/7 incident response readiness given baseline threat levels
- Continue federal-state collaboration across intelligence, infrastructure, and tech

---

# Appendix

```{r loss-summary}
# Summary of losses by attack type
cyber_data %>%
  group_by(attack_type) %>%
  summarize(
    total_loss = sum(estimated_loss_usd),
    avg_loss = mean(estimated_loss_usd),
    .groups = 'drop'
  ) %>%
  arrange(desc(total_loss)) %>%
  kable(digits = 2)
```